Short-time expansion of Heisenberg operators in open collective quantum spin systems

### 2019-09-13 soc_squeezing/collective_paper/collective_spin.tex
pushing submitted version of collective spin paper

%%% figures
\usepackage{graphicx} % for figures
\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
\newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures
\graphicspath{{./figures/}} % set path for all figures

@@

Collective spin systems are a versatile resource in quantum science
for a range of applications including quantum-enhanced metrology and
quantum simulation.  The study of such systems dates back to the
mid-twentieth century with the introduction of the Dicke
model\cite{dicke1954coherence} that describes atoms cooperatively
interacting with a single mode of a radiation field, and the
Lipkin-Meshkov-Glick (LMG) model, a toy model for testing many-body
approximation methods in contemporary nuclear
physics\cite{lipkin1965validity, meshkov1965validity,
  glick1965validity}.  On the experimental side, the development of
advanced trapping, cooling, and control techniques in atomic,
molecular, and optical (AMO) systems have enabled the realization of
collective spin models in a broad range of platforms, including atomic
ensembles\cite{takano2009spin, appel2009mesoscopic,
  schleier-smith2010states, chen2011conditional}, Bose-Einstein
condensates\cite{esteve2008squeezing, riedel2010atomchipbased,
  gross2010nonlinear, klinder2015dynamical}, ultracold Fermi
gasses\cite{martin2013quantum}, trapped ions\cite{bohnet2016quantum},
and optical cavities\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement,
  norcia2018cavitymediated}, among others.  These implementations
compliment theoretical studies in a variety of rich subjects,
including quantum phase transitions and
criticality\cite{latorre2005entanglement, alcalde2007functional,
  wang2012quantum, majd2014lmg}, non-equilibrium
phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
  morrison2008dissipationdriven, morrison2008collective,
  kessler2012dissipative, bhattacherjee2014nonequilibrium,
  zhiqiang2017nonequilibrium, lang2018concurrence}, and precision
mentrology\cite{wineland1992spin, kitagawa1993squeezed,
  zhong2010simplified, schleier-smith2010squeezing, ma2011quantum,
  huang2015twoaxis, muessel2015twistandturn, huang2015quantum,
  mirkhalaf2018robustifying, lewis-swan2018robust, he2019engineering}.

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum}.
Collective spin systems provide a means to break through this limit
via the preparation of many-body entangled states such as spin-cat
states\cite{agarwal1997atomic, lau2014proposal, huang2015quantum} and
most notably spin-squeezed states\cite{wineland1992spin,
  kitagawa1993squeezed, ma2011quantum} that allow for measurement
errors $\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$,
where $\varepsilon=1$ saturates the Heisenberg
limit\cite{zwierz2010general}.  Such entangled states can be prepared
either via heralded methods such as quantum nondemolition
measurements\cite{takano2009spin, appel2009mesoscopic,
  schleier-smith2010states, chen2011conditional}, or via deterministic
methods that require nonlinear dynamics, typically realized with
phonon-mediated\cite{bohnet2016quantum},
photon-mediated\cite{leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement,
  norcia2018cavitymediated} or collisional\cite{esteve2008squeezing,
  riedel2010atomchipbased, gross2010nonlinear, bromley2018dynamics}
interactions.  Although a truly collective spin model requires
uniform, all-to-all interactions, as long as measurements do not
distinguish between constituent particles, even non-uniform systems
may be effectively described by a uniform model with renormalized
parameters\cite{hu2015entangled}.

@@

In this work, we present a new method to compute short-time dynamics
of collective spin systems with generic Markovian decoherence.  The
only restriction on decoherence (beyond Markovianity) is that, like
the coherent collective dynamics, it must act identically on all
constituent particles.  Our method is based on a short-time expansion
of exact solutions to the equations of motion for Heisenberg
operators.  Evaluating this expansion requires knowing the structure
constants of a collective spin operator algebra; the calculation of
these structure constants (in Appendices
\ref{sec:identities}--\ref{sec:general_product}) is one of the main
technical results of this work, which we hope will empower both
analytical and numerical studies of collective spin systems in the
future.  We benchmark our method against both exact and quantum
trajectory Monte Carlo computations of spin squeezing in accessible
parameter regimes, highlighting both advantages and limitations of our
short-time expansion.  Finally, we showcase applications of our method
by computing quantities that are inaccessible to other numerical
methods.

@@

\input{calculations_main.tex}

\input{tutorial.tex}

\input{calculations_OAT.tex}

\input{noise_operators.tex}

\input{linear_response.tex}

% \input{changing_bases.tex}

>>>>

%%% figures
\usepackage{graphicx} % for figures
\usepackage[caption=false]{subfig} % subfigures (via \subfloat[]{})
\newcommand{\sref}[1]{\protect\subref{#1}} % for referencing subfigures

@@

Collective spin systems are a versatile resource in quantum science
for a range of applications including quantum-enhanced metrology and
quantum simulation.  The study of such systems dates back to the
mid-twentieth century with the introduction of the Dicke
model\cite{dicke1954coherence} that describes atoms cooperatively
interacting with a single mode of a radiation field, and the
Lipkin-Meshkov-Glick (LMG) model, a toy model for testing many-body
approximation methods in contemporary nuclear
physics\cite{lipkin1965validity, meshkov1965validity,
  glick1965validity}.  On the experimental side, the development of
advanced trapping, cooling, and control techniques in atomic,
molecular, and optical (AMO) systems have enabled the realization of
collective spin models in a broad range of platforms, including cold
atomic gasses\cite{takano2009spin, appel2009mesoscopic}, Bose-Einstein
condensates\cite{klinder2015dynamical, esteve2008squeezing,
  riedel2010atomchipbased, gross2010nonlinear}, ultracold Fermi
gasses\cite{martin2013quantum, bromley2018dynamics,
  smale2019observation}, trapped ions\cite{bohnet2016quantum}, and
optical cavities\cite{schleier-smith2010states, chen2011conditional,
  baumann2010dicke, leroux2010implementation, bohnet2014reduced,
  cox2016deterministic, hosten2016measurement, hosten2016quantum,
  norcia2018cavitymediated, ritsch2013cold}, among others.  These
implementations compliment innumerable theoretical studies in a
variety of rich subjects, including quantum phase transitions and
criticality\cite{latorre2005entanglement, alcalde2007functional,
  wang2012quantum, majd2014lmg}, non-equilibrium
phenomena\cite{walls1978nonequilibrium, morrison2008dynamical,
  morrison2008dissipationdriven, morrison2008collective,
  kessler2012dissipative, bhattacherjee2014nonequilibrium,
  zhiqiang2017nonequilibrium, lang2018concurrence}, and precision
mentrology\cite{wineland1992spin, kitagawa1993squeezed,
  zhong2010simplified, schleier-smith2010squeezing, ma2011quantum,
  huang2015twoaxis, muessel2015twistandturn, huang2015quantum,
  hu2017vacuum, mirkhalaf2018robustifying, lewis-swan2018robust,
  he2019engineering}.

One of the primary motivations for studying collective spin systems is
their application to quantum-enhanced metrology.  Quantum projection
noise limits the error $\Delta\phi$ in the measurement of a phase
angle $\phi$ with $N$ independent spins to
$\Delta\phi\sim1/\sqrt{N}$\cite{wineland1992spin, itano1993quantum}.
Collective spin systems provide a means to break through this limit
via the preparation of many-body entangled states such as spin-cat
states\cite{agarwal1997atomic, lau2014proposal, huang2015quantum} and
most notably spin-squeezed states\cite{wineland1992spin,
  kitagawa1993squeezed, ma2011quantum} that allow for measurement
errors $\Delta\phi\sim1/N^\varepsilon$ with $1/2<\varepsilon\le1$,
where $\varepsilon=1$ saturates the Heisenberg
limit\cite{zwierz2010general}.  Such entangled states can be prepared
either via heralded methods such as quantum nondemolition
measurements\cite{takano2009spin, appel2009mesoscopic,
  schleier-smith2010states, chen2011conditional}, or via deterministic
methods that require nonlinear dynamics, typically realized with
phonon-mediated\cite{bohnet2016quantum},
photon-mediated\cite{klinder2015dynamical, baumann2010dicke,
  leroux2010implementation, bohnet2014reduced, cox2016deterministic,
  hosten2016measurement, hosten2016quantum, norcia2018cavitymediated,
  ritsch2013cold} or collisional\cite{esteve2008squeezing,
  riedel2010atomchipbased, gross2010nonlinear, martin2013quantum,
  bromley2018dynamics, smale2019observation} interactions.  Although a
truly collective spin model requires uniform, all-to-all interactions,
as long as measurements do not distinguish between constituent
particles, even non-uniform systems may be effectively described by a
uniform model with renormalized parameters\cite{hu2015entangled}.

@@

In this work, we present a new method to compute short-time dynamics
of collective spin systems with generic Markovian decoherence.  The
only restriction on decoherence (beyond Markovianity) is that, like
the coherent collective dynamics, it must act identically on all
constituent particles.  Our method is based on a short-time expansion
of exact solutions to the equations of motion for Heisenberg
operators.  Evaluating this expansion requires knowing the structure
constants of a collective spin operator algebra; the calculation of
these structure constants (in Appendices
\ref{sec:identities}--\ref{sec:general_product}) is one of the main
technical results of this work, which we hope will empower both
analytical and numerical studies of collective spin systems in the
future.  We benchmark our method against exact results from both
analytical calculations and quantum trajectory Monte Carlo
computations of spin squeezing in accessible parameter regimes,
highlighting both advantages and limitations of the short-time
expansion.  Finally, we showcase applications of our method by
computing quantities that are inaccessible to other numerical methods.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Basic spin operator identities}
\label{sec:identities}

The appendices in this work make ubiquitous use of various spin
operator identities; we collect and derive some basic identities here
for reference.  Note that despite the working definition of collective
spin operators from $S_\a=\sum_js_\a^{(j)}$, the identities we will
derive involving only collective spin operators apply just as well to
large-spin operators that cannot be expressed as the sum of individual
spin-1/2 operators.  The elementary commutation relations between spin
operators are, with $\bmu\equiv-\mu\in\set{+1,-1}$ for brevity,
\begin{align}
  \sp{s_\z^{(j)},s_\mu^{(k)}}_-
  &= \delta_{jk} \mu s_\mu^{(j)},
  &
  \sp{S_\z,s_\mu^{(j)}}_-
  &= \sp{s_\z^{(j)},S_\mu}_- = \mu s_\mu^{(j)},
  &
  \sp{S_\z,S_\mu}_-
  &= \mu S_\mu,
  \label{eq:comm_z_base} \\
  \sp{s_\mu^{(j)},s_\bmu^{(k)}}_-
  &= \delta_{jk} 2 \mu s_\z^{(j)},
  &
  \sp{S_\mu,s_\bmu^{(j)}}_-
  &= \sp{s_\mu^{(j)},S_\bmu}_- = 2 \mu s_\z^{(j)},
  &
  \sp{S_\mu,S_\bmu}_-
  &= 2 \mu S_\z.
  \label{eq:comm_mu_base}
\end{align}
These relations can be used to inductively compute identities
involving powers of collective spin operators.  By pushing through one
spin operator at a time, we can find
\begin{align}
  \p{\mu S_\z}^m s_\mu^{(j)}
  = \p{\mu S_\z}^{m-1} s_\mu^{(j)} \p{1 + \mu S_\z}
  = \p{\mu S_\z}^{m-2} s_\mu^{(j)} \p{1 + \mu S_\z}^2
  = \cdots
  = s_\mu^{(j)} \p{1 + \mu S_\z}^m,
  \label{eq:push_z_mu_Ss}
\end{align}
and
\begin{align}
  \mu s_\z^{(j)} S_\mu^m
  = S_\mu \mu s_\z^{(j)} S_\mu^{m-1} + s_\mu^{(j)} S_\mu^{m-1}
  = \cdots
  = S_\mu^m \mu s_\z^{(j)} + ms_\mu^{(j)} S_\mu^{m-1},
  \label{eq:push_z_mu_sS}
\end{align}
where we will generally find it nicer to express results in terms of
$\mu s_\z^{(j)}$ and $\mu S_\z$ rather than $s_\z^{(j)}$ and $S_\z$.
Summing over the single-spin index $j$ in both of the cases above
gives us the purely collective-spin versions of these identities:
\begin{align}
  \p{\mu S_\z}^m S_\mu = S_\mu \p{1 + \mu S_\z}^m,
  &&
  \mu S_\z S_\mu^m = S_\mu^m \p{m + \mu S_\z},
  \label{eq:push_z_mu_single}
\end{align}
where we can repeat the process of pushing through individual $S_\z$
operators $\ell$ times to get
\begin{align}
  \p{\mu S_\z}^\ell S_\mu^m
  = \p{\mu S_\z}^{\ell-1} S_\mu^m \p{m + \mu S_\z}
  = \p{\mu S_\z}^{\ell-2} S_\mu^m \p{m + \mu S_\z}^2
  = \cdots
  = S_\mu^m \p{m + \mu S_\z}^\ell.
  \label{eq:push_z_mu}
\end{align}
Multiplying \eqref{eq:push_z_mu} through by $\p{\mu\nu}^\ell$ (for
$\nu\in\set{+1,-1}$) and taking its Hermitian conjugate, we can say
that more generally
\begin{align}
  \p{\nu S_\z}^\ell S_\mu^m
  = S_\mu^m \p{\mu\nu m+\nu S_\z}^\ell,
  &&
  S_\mu^m \p{\nu S_\z}^\ell
  = \p{-\mu\nu m+\nu S_\z}^\ell S_\mu^m.
\end{align}
Finding commutation relations between powers of transverse spin
operators, i.e.~$S_\mu$ and $S_\bmu$, turns out to be considerably
more difficult than the cases we have worked out thus far.  We
therefore save this work for Appendix \ref{sec:comm_transverse}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Commutation relations between powers of transverse spin
  operators}
\label{sec:comm_transverse}

To find commutation relations between powers of transverse collective
spin operators, we first compute
\begin{align}
  S_\mu^m s_\bmu^{(j)}
  &= S_\mu^{m-1}s_\bmu^{(j)} S_\mu
  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
  &= S_\mu^{m-2} s_\bmu^{(j)} S_\mu^2
  + S_\mu^{m-2} 2\mu s_\z^{(j)} S_\mu
  + S_\mu^{m-1} 2\mu s_\z^{(j)} \\
  &= s_\bmu^{(j)} S_\mu^m
  + \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
  \label{eq:push_mu_Ss_start}.
\end{align}
While \eqref{eq:push_mu_Ss_start} gives us the commutator
$\sp{S_\mu^m,s_\bmu^{(j)}}_-$, we would like to enforce an ordering on
products of spin operators, which will ensure that we only keep track
of operators that are linearly independent.  We choose (for now) to
impose an ordering with all $s_\bmu^{(j)}$ operators on the left, and
all $s_\z^{(j)}$ operators on the right.  Such an ordering will prove
convenient for the calculations in this section\footnote{In
  retrospect, it may have been nicer to push all $s_\mu^{(j)}$
  operators to the right throughout these calculations, due to the
  enhanced symmetry that expressions would have with respect to
  Hermitian conjugation.  In any case, we provide the final result of
  this section in both ordering conventions, and therefore feel no
  need to reproduce these calculations with a different ordering of
  spin operators.}.  This choice of ordering compels us to expand
\begin{align}
  \sum_{k=0}^{m-1} S_\mu^k 2\mu s_\z^{(j)} S_\mu^{m-k-1}
  &= \sum_{k=0}^{m-1} S_\mu^k
  \sp{2\p{m-k-1} s_\mu^{(j)} S_\mu^{m-k-2}
    + S_\mu^{m-k-1} 2\mu s_\z^{(j)}} \\
  &= m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
\end{align}
which implies
\begin{align}
  S_\mu^m s_\bmu^{(j)}
  = s_\bmu^{(j)} S_\mu^m + m \p{m-1} s_\mu^{(j)} S_\mu^{m-2}
  + m S_\mu^{m-1} 2\mu s_\z^{(j)},
  \label{eq:push_mu_Ss}
\end{align}
and in turn
\begin{align}
  S_\mu^m S_\bmu = S_\bmu S_\mu^m
  + m S_\mu^{m-1} \p{m - 1 + 2\mu S_\z}.
  \label{eq:push_mu_single}
\end{align}
As the next logical step, we take on the task of computing
\begin{align}
  S_\mu^m S_\bmu^n
  = S_\mu^{m-1} S_\bmu^n S_\mu
  + n \sp{S_\mu^{m-1} S_\bmu^{n-1} \p{1 - n + 2\mu S_\z}}
  = S_\bmu^n S_\mu^m
  + n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
  \p{1 - n + 2\mu S_\z} S_\mu^k,
\end{align}
which implies
\begin{align}
  \sp{S_\mu^m, S_\bmu^n}_-
  = C_{mn;\mu}
  \equiv n \sum_{k=0}^{m-1} S_\mu^{m-k-1} S_\bmu^{n-1}
  \p{1 - n + 2\mu S_\z} S_\mu^k.
\end{align}
We now need rearrange the operators in $C_{mn;\mu}$ into a standard
order, which means pushing all $S_\z$ operators to the right and, for
the purposes of this calculation, all $S_\bmu$ operators to the left.
We begin by pushing $S_\mu^k$ to the left of $S_\z$, which takes
$2\mu S_\z\to 2\mu S_\z+2k$, and then push $S_\mu^{m-k-1}$ to the
right of $S_\bmu^{n-1}$, giving us
\begin{align}
  C_{mn;\mu}
  &= n \sum_{k=0}^{m-1}
  \p{S_\bmu^{n-1} S_\mu^{m-k-1} + C_{m-k-1,n-1;\mu}} S_\mu^k
  \p{2k + 1 - n + 2\mu S_\z} \\
  &= D_{mn;\mu}
  + n \sum_{k=0}^{m-2} C_{m-k-1,n-1;\mu}
  S_\mu^k \p{2k + 1 - n + 2\mu S_\z},
  \label{eq:C_mn}
\end{align}
where we have dropped the last ($k=m-1$) term in the remaining sum
because $C_{m-k-1,n-1;\mu}=0$ if $k=m-1$, and
\begin{align}
  D_{mn;\mu}
  \equiv mn S_\bmu^{n-1} S_\mu^{m-1} \p{m - n + 2\mu S_\z}.
  \label{eq:D_mn}
\end{align}
To our despair, we have arrived in \eqref{eq:C_mn} at a {\it
  recursive} formula for $C_{mn;\mu}$.  Furthermore, we have not even
managed to order all spin operators, as $C_{m-k-1,n-1;\mu}$ contains
$S_\z$ operators that are to the left of $S_\mu^k$.  To sort all spin
operators once and for all, we define
\begin{align}
  C_{mn;\mu}^{(k)} \equiv C_{m-k,n;\mu} S_\mu^k,
  &&
  D_{mn;\mu}^{(k)} \equiv D_{m-k,n;\mu} S_\mu^k,
\end{align}
which we can expand as
\begin{align}
  D_{mn;\mu}^{(k)}
  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-k-1}
  \p{m-k-n+2\mu S_\z} S_\mu^k \\
  &= \p{m-k}n S_\bmu^{n-1} S_\mu^{m-1} \p{k+m-n+2\mu S_\z},
  \label{eq:D_mn_k}
\end{align}
and
\begin{align}
  C_{mn;\mu}^{(k)}
  &= D_{m-k,n;\mu} S_\mu^k + n \sum_{j=0}^{m-k-2}
  C_{m-k-j-1,n-1;\mu} S_\mu^j \p{2j+1-n+2\mu S_\z} S_\mu^k \\
  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
  C_{m-k-j-1,n-1;\mu} S_\mu^{j+k} \p{2j+2k+1-n+2\mu S_\z} \\
  &= D_{mn;\mu}^{(k)} + n \sum_{j=0}^{m-k-2}
  C_{m-1,n-1;\mu}^{(k+j)} \p{2\sp{j+k}+1-n+2\mu S_\z} \\
  &= D_{mn;\mu}^{(k)} + n \sum_{j=k}^{m-2}
  C_{m-1,n-1;\mu}^{(j)} \p{2j+1-n+2\mu S_\z}.
  \label{eq:C_mn_k}
\end{align}
While the resulting expression in \eqref{eq:C_mn_k} strongly resembles
that in \eqref{eq:C_mn}, there is one crucial difference: all spin
operators in \eqref{eq:C_mn_k} have been sorted into a standard order.
We can now repeatedly substitute $C_{mn;\mu}^{(k)}$ into itself, each
time decreasing $m$ and $n$ by 1, until one of $m$ or $n$ reaches
zero.  Such repeated substitution yields the expansion
\begin{align}
  C_{mn;\mu}
  = C_{mn;\mu}^{(0)}
  = D_{mn;\mu}
  + \sum_{p=1}^{\min\set{m,n}-1} E_{mn;\mu}^{(p)},
  \label{eq:C_mn_E}
\end{align}
where the first two terms in the sum over $p$ are
\begin{align}
  E_{mn;\mu}^{(1)}
  &= n \sum_{k=0}^{m-2} D_{m-1,n-1;\mu}^{(k)} \p{2k+1-n+2\mu S_\z}, \\
  E_{mn;\mu}^{(2)}
  &= n \sum_{k_1=0}^{m-2} \p{n-1} \sum_{k_2=k_1}^{m-3}
  D_{m-2,n-2;\mu}^{(k_2)} \p{2k_2+2-n+2\mu S_\z} \p{2k_1+1-n+2\mu S_\z},
\end{align}
and more generally for $p>1$,
\begin{align}
  E_{mn;\mu}^{(p)}
  = \f{n!}{\p{n-p}!}
  \sum_{k_1=0}^{m-2} \sum_{k_2=k_1}^{m-3} \cdots\sum_{k_p=k_{p-1}}^{m-p-1}
  D_{m-p,n-p;\mu}^{(k_p)} \prod_{j=1}^p \p{2k_j+j-n+2\mu S_\z}.
  \label{eq:E_mn_p}
\end{align}
In principle, the expressions in \eqref{eq:D_mn}, \eqref{eq:D_mn_k},
\eqref{eq:C_mn_E}, and \eqref{eq:E_mn_p} suffice to evaluate the
commutator $\sp{S_\mu^m,S_\bmu^n}_- = C_{mn;\mu}$, but this result is
-- put lightly -- quite a mess: the expression for $E_{mn;\mu}^{(p)}$
in \eqref{eq:E_mn_p} involves a sum over $p$ mutually dependent
intermediate variables, each term of which additionally contains a
product of $p$ factors.  We therefore devote the rest of this section
to simplifying our result for the commutator
$\sp{S_\mu^m,S_\bmu^n}_-$.

Observing that in \eqref{eq:E_mn_p} we always have
$0\le k_1\le k_2\le\cdots\le k_p\le m-p-1$, we can rearrange the order
of the sums and relabel $k_p\to\ell$ to get
\begin{align}
  E_{mn;\mu}^{(p)}
  = \f{n!}{\p{n-p}!}
  \sum_{\ell=0}^{m-p-1} D_{m-p,n-p;\mu}^{(\ell)} \p{2\ell+F_{np;\mu}}
  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}},
  \label{eq:E_mn_p_sum}
\end{align}
where for shorthand we define
\begin{align}
  F_{np;\mu} \equiv p - n + 2\mu S_\z,
  &&
  \sum_{\p{\v k,q,\ell}} X \equiv
  \sum_{k_1=0}^\ell \sum_{k_2=k_1}^\ell
  \cdots \sum_{k_q=k_{q-1}}^\ell X.
\end{align}
We now further define
\begin{align}
  f_{np\ell;\mu}\p{k,q} \equiv \p{\ell-k+q} \p{\ell+k-q+F_{np;\mu}},
\end{align}
and evaluate sums successively over $k_{p-1},k_{p-2},\cdots,k_1$,
finding
\begin{align}
  \sum_{\p{\v k,p-1,\ell}} \prod_{j=1}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  &= \sum_{\p{\v k,p-2,\ell}}
  \prod_{j=2}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  f_{np\ell;\mu}\p{k_{p-2},1} \\
  &= \f1{\p{r-1}!} \sum_{\p{\v k,p-r,\ell}}
  \prod_{j=r}^{p-1} \p{2k_{p-j}-j+F_{np;\mu}}
  \prod_{q=1}^{r-1} f_{np\ell;\mu}\p{k_{p-r},q} \\
  &= \f1{\p{p-1}!} \prod_{q=1}^{p-1} f_{np\ell;\mu}\p{0,q} \\
  &= { \ell + p - 1 \choose p - 1 }
  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}}.
\end{align}
Substitution of this result together with $D_{m-p,n-p;\mu}^{(\ell)}$
using \eqref{eq:D_mn_k} into \eqref{eq:E_mn_p_sum} then gives us
\begin{align}
  E_{mn;\mu}^{(p)}
  = \f{n!}{\p{n-p-1}!} S_\bmu^{n-p-1} S_\mu^{m-p-1} G_{mnp;\mu}
\end{align}
with
\begin{align}
  G_{mnp;\mu}
  &\equiv \sum_{\ell=0}^{m-p-1} { \ell + p - 1 \choose p - 1 }
  \p{m-p-\ell} \p{\ell+m-p+F_{np;\mu}}
  \p{2\ell + F_{np;\mu}}
  \prod_{q=1}^{p-1} \p{\ell-q+F_{np;\mu}} \\
  &= { m \choose p + 1 } \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}.
\end{align}
We can further simplify
\begin{align}
  \prod_{q=0}^p \p{m-p-q+F_{np;\mu}}
  = \prod_{q=0}^p \p{m-n-q+2\mu S_\z}
  = \sum_{q=0}^{p+1} \p{-1}^{p+1-q}
  { p+1 \brack q } \p{m-n+2\mu S_\z}^q,
\end{align}
where ${ p \brack q }$ is an unsigned Stirling number of the first
kind, and finally
\begin{align}
  \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \p{m-n+2\mu S_\z}^q
  &= \sum_{q=0}^p \p{-1}^{p-q} { p \brack q } \sum_{\ell=0}^q
  { q \choose \ell } \p{m-n}^{q-\ell} \p{2\mu S_\z}^\ell \\
  &= \sum_{\ell=0}^p 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell} \p{\mu S_\z}^\ell.
\end{align}
Putting everything together, we finally have
\begin{align}
  E_{mn;\mu}^{(p-1)}
  = p! { m \choose p } { n \choose p }
  S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
\end{align}
with
\begin{align}
  \epsilon_{mn}^{p\ell}
  \equiv 2^\ell \sum_{q=\ell}^p \p{-1}^{p-q}
  { p \brack q } { q \choose \ell } \p{m-n}^{q-\ell},
\end{align}
where in this final form $E_{mn;\mu}^{(0)} = D_{mn;\mu}$, which
together with the expansion for $C_{mn;\mu}$ in \eqref{eq:C_mn_E}
implies that
\begin{align}
  \sp{S_\mu^m, S_\bmu^n}_-
  = \sum_{p=1}^{\min\set{m,n}}
  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell,
  \label{eq:comm_mu}
\end{align}
and
\begin{align}
  S_\mu^m S_\bmu^n
  = \sum_{p=0}^{\min\set{m,n}}
  p! { m \choose p } { n \choose p } S_\bmu^{n-p} S_\mu^{m-p}
  \sum_{\ell=0}^p \epsilon_{mn}^{p\ell} \p{\mu S_\z}^\ell.
  \label{eq:push_mu_bmu}
\end{align}
If we wish to order products of collective spin operators with $S_\z$
in between $S_\bmu$ and $S_\mu$, then
\begin{align}
  S_\mu^m S_\bmu^n
  = \sum_{p=0}^{\min\set{m,n}} p! { m \choose p } { n \choose p }
  S_\bmu^{n-p} Z_{mn;\bmu}^{(p)} S_\mu^{m-p},
\end{align}
where
\begin{align}
  Z_{mn;\bmu}^{(p)}
  \equiv \sum_{\ell=0}^p \epsilon_{mn}^{p\ell}
  \p{-\sp{m-p} + \mu S_\z}^\ell
  = \sum_{q=0}^p \zeta_{mn}^{pq} \p{\bmu S_\z}^q,
  \label{eq:Z_mnp}
\end{align}
with
\begin{align}
  \zeta_{mn}^{pq}
  \equiv \sum_{\ell=q}^p \epsilon_{mn}^{p\ell}
  { \ell \choose q } \p{-1}^\ell \p{m-p}^{\ell-q}
  = \p{-1}^p 2^q \sum_{s=q}^p
  { p \brack s } { s \choose q } \p{m+n-2p}^{s-q}.
  \label{eq:zeta_mnpq}
\end{align}
Here ${ p \brack s }$ is an unsigned Stirling number of the first
kind.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Product of arbitrary ordered collective spin operators}
\label{sec:general_product}

The most general product of collective spin operators that we need to
compute is
\begin{align}
  \S^{pqr}_{\ell mn;\mu}
  = S_\mu^p \p{\mu S_\z}^q S_\bmu^r
  S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n
  = \sum_{k=0}^{\min\set{r,\ell}} k! { r \choose k } { \ell \choose k }
  S_\mu^{p+\ell-k} \tilde Z_{qr\ell m;\mu}^{(k)} S_\bmu^{r+n-k},
  \label{eq:general_product}
\end{align}
where
\begin{align}
  \tilde Z_{qr\ell m;\mu}^{(k)}
  &\equiv \p{\ell-k+\mu S_\z}^q
  Z_{r\ell;\mu}^{(k)} \p{r-k+\mu S_\z}^m \\
  &= \sum_{a=0}^k \zeta_{r\ell}^{ka}
  \sum_{b=0}^q \p{\ell-k}^{q-b} { q \choose b }
  \sum_{c=0}^m \p{r-k}^{m-c} { m \choose c }
  \p{\mu S_\z}^{a+b+c},
\end{align}
is defined in terms of $Z_{r\ell;\mu}^{(k)}$ and $\zeta_{r\ell}^{ka}$
as respectively given in \eqref{eq:Z_mnp} and \eqref{eq:zeta_mnpq}.
The (anti-)commutator of two ordered products of collective spin
operators is then simply
\begin{align}
  \sp{S_\mu^p \p{\mu S_\z}^q S_\bmu^r,
    S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n}_\pm
  = \S^{pqr}_{\ell mn;\mu} \pm \S^{\ell mn}_{pqr;\mu}.
\end{align}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Sandwich identities for single-spin decoherence calculations}
\label{sec:sandwich_single}

In this section we derive several identities that will be necessary
for computing the effects of single-spin decoherence on ordered
products of collective spin operators, i.e.~on operators of the form
$S_\mu^\ell \p{\mu S_\z}^m S_\bmu^n$.  These identities all involve
sandwiching a collective spin operator between operators that act on
individual spins only, and summing over all individual spin indices.
Our general strategy will be to use commutation relations to push
single-spin operators together, and then evaluate the sum to arrive at
an expression involving only collective spin operators.
